---
title: "Forest Growth - ODE and Sobel Sensitivity"
author: "Anna Abelman, Margaret Brickner, & Hannah Garcia"
date: "5/6/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
# Load packages 
library(tidyverse)
library(purrr)
library(deSolve)
library(sensitivity)
library(tinytex)
```

```{r}
# read in function
source("forest_growth.R")
```

### ODE for 300 years
Run the model for 300 years (using the ODE solver)  starting with an initial forest size of 10 kg/C, and using the following parameters: canopy closure threshold of 50 kgC , K = 250 kg C (carrying capacity) , r=  0.01 (exponential growth rate before before canopy closure), g = 2 kg/year (linear growth rate after canopy closure)
```{r}
#set all of the parameters so they're easier to update and easier to track. 
forest_parms = list(K=250, r = 0.01, g = 2, Ct = 50)

#set the years and initial size
years = seq(from=1, to=300)
initial_size = 10

#use ODE solver to run the model for 300 years
forest_300_result = ode(y = initial_size, times = years, func = forest_growth, parms = forest_parms)

#check out the results
colnames(forest_300_result)=c("year","size")
head(forest_300_result)

#graph the Results
ggplot(as.data.frame(forest_300_result), aes(year, size))+
  geom_point(colour = "forestgreen", size = .3) + 
  labs(title = "Modeled Forest Growth", x = "Year", y = "Forest Size (kg/C)")+ 
  theme_minimal() +
  theme(axis.line = element_line(color = "dimgrey"), 
        panel.border = element_rect(fill = NA, color = NA), 
        panel.grid.minor.y = element_blank(), 
        panel.grid.minor.x = element_blank()) + 
  scale_x_continuous(expand = c(0.0,1)) + 
  scale_y_continuous(expand = c(0.0,1))

```


### Sobel Sensitivity
Run a sobol sensitivity analysis that explores how the estimated maximum and mean forest size (e.g maximum and mean values of C over the 300 years) varies with the pre canopy closure growth rate (r) and post-canopy closure growth rate (g) and canopy closure threshold and carrying capacity(K)

Assume that they are all normally distributed with means as given above and standard deviation of 10% of mean value
```{r}
#initial population value
Pinitial=10

#remembering the initial parameters
#forest_parms = list(K=250, r = 0.01, g = 2, Ct = 50)

#first set the number of parameters
np=100
K = rnorm(mean=250, sd=25, n=np)
r = rnorm(mean=0.01, sd=0.001, n=np)
g = rnorm(mean=2, sd=0.2, n=np)
X1 = cbind.data.frame(r=r, K=K, g=g, Ct=50)

#second set of samples
K = rnorm(mean=250, sd=25, n=np)
r = rnorm(mean=0.01, sd=0.001, n=np)
g = rnorm(mean=2, sd=0.2, n=np)
X2 = cbind.data.frame(r=r, K=K, g=g, Ct=50)

#create sobel object that holds the 2 sample datasets
sens_parms = soboljansen(model = NULL, X1, X2, nboot = 300)

#checking out the first values
head(sens_parms$X)

# gets results for each year for 300 years 
times = seq(from=1, to=300) #number of simulations runs
parms = list(r=sens_parms$X$r[1], K=sens_parms$X$K[1], g=sens_parms$X$g[1], Ct=sens_parms$X$Ct[1]) #first set of parameters
forest_result = ode(y=Pinitial, times=times, func=forest_growth, parms=parms) #run ODE to get results for first set

#checking out first values
head(forest_result)
colnames(forest_result)=c("time","P") #change column names so easier to use

#change to a dataframe to plot
forest_result = as.data.frame(forest_result)
ggplot(forest_result, aes(time, P))+
  geom_point()

#this plot confirms our ODE results make sense so now we can begin running with the other parameters to see how sensitive our analysis is based on the parameters: g, K, and r
```

### Sensitivity Analysis of Maximum Forest Size
Graph the results of the sensitivity analysis as a box plot of maximum forest size and a plot of the two sobol indices (S and T)

```{r}
#max population from first set of parameters
max_pop = max(forest_result$P)
max_pop

#how many years required to get to the maximum poputation
max_year = which.max(forest_result$P)
```

```{r, warning=FALSE, message=FALSE, results='hide'}
#create a function to pull out max_year and max_pop
compute_max_metrics = function(forest_result) {
  max_pop = max(forest_result$P)
  idx = which.max(forest_result$P)
  max_year = forest_result$time[idx]
return(list(max_pop=max_pop, max_year=max_year))
}

#try it on our first parameter set
compute_max_metrics(forest_result)

#create a wrapper function that will solve and return the max_pop and max_year for each set of parameters created above
forest_wrapper = function(r, K, g, Ct, Pinitial, times, func) {
    parms = list(r=r, K=K, g=g, Ct=Ct) #list parameters needed
    forest_result = ode(y=Pinitial, times=times, func=forest_growth, parms=parms) #solve function
    colnames(forest_result)=c("time","P") #change column names
  max_metrics=compute_max_metrics(as.data.frame(forest_result)) #find max_pop and max_year
  return(max_metrics) 
}

#use pmap tp return all results for all 100 sets of parameters
all_forest_results = sens_parms$X %>% 
  pmap(forest_wrapper, Pinitial=Pinitial, times=times, func=forest_growth)

#use results from above (pmap) and turn into nice dataframe
all_forest_results_df = all_forest_results %>% map_dfr(`[`,c("max_pop","max_year"))


#turn into tidy format
final_forest = all_forest_results_df %>% 
  gather(key="metric", value="value")
```

```{r}
#plot it!
ggplot(final_forest, aes(metric, value, col=metric))+
  geom_boxplot()+
  theme_classic()
```

Looking at the sensitivity for each parameter on `max_pop`
```{r}
sens_parms_max_pop = sensitivity::tell(sens_parms,all_forest_results_df$max_pop)
#first-order indices(S)
sens_parms_max_pop$S
#total sensitivity index (T)
sens_parms_max_pop$T
#plot to see the difference
plot(sens_parms_max_pop)
```

Looking at the sensitivity for each parameter on `max_year`
```{r}
sens_parms_max_year = sensitivity::tell(sens_parms,all_forest_results_df$max_year)
#first-order indices (S)
sens_parms_max_year$S
#total sensitivity index (T)
sens_parms_max_year$T
#plot to see the difference
plot(sens_parms_max_year)
```

